Everything you wanted to know about PanDA but were truly truly afraid to ask

### 2017-03-11 overall.tex
Section V ready for comments

Currently, 20 instances of PanDA Broker are deployed on a total of 4 DTNs, 5
instances for each DTN. Each broker controls from 15 to 300 detector simulation
jobs per submission, one simulation job for each Titan's work node and a
theoretical maximum concurrent use of 96,000 cores. Since November 2015, the
PanDA Brokers have operated only in backfill mode, without a defined time
allocation, and running at the lowest priority on Titan.

We evaluated the efficiency, scalability and reliability of the PanDA Brokers
deployed on Titan along two dimensions: (i) amount of backfill availability
utilized by the PanDA Broker; and (ii) amount of jobs walltime spent performing
detector simulations. We based these evaluations on measuring the number of
cores utilized by ATLAS on Titan, the overall backfill availability on Titan,
the number of detector simulations performed on Titan and on all the resources
available to ATLAS, and the failure rate of these simulations. All the
measurements were performed between January 2016 and February 2017.

@@

On February 2017, PanDA Brokers used almost twice as much backfill availability
than in any other month. No relevant code update was made during that period and
logs indicated that the brokers were able to respond more promptly to backfill
availability. This is likely due to hardware upgrades on the DTNs. The absence
of continuous monitoring of the DTNs does not allow to quantify bottlenecks but
spot measurements of the DTNs load indicate that a faster CPU and better
networking were likely responsible for the improved performance.

Investigations of the CPU load of the upgraded DTNs shows an average
XX\mtnote{asked Danila for data} load. As such, further hardware upgrades seem
unlikely to improve significantly the performance of PanDA Brokers. Nonetheless,
the current load suggests that the number of brokers per DTN could be increased.
This would enable the submission of a larger number of concurrent jobs to
Titan's PBS queue, allowing for PanDA to consume a higher percentage of
backfill availability.

Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency from
January 2016 to February 2017. Efficiency is defined as the fraction of Titan’s
cores available via backfill that was utilized by the PanDA Brokers. The brokers
reached 18\% average efficiency with a minimum 8.9\% efficiency on April 2016
and a maximum 33.5\% efficiency on February 2017. The number of total backfill
cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
that the difference in total backfill availability was not responsible for the measured difference in efficiency.

@@

Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
February 2017, about 2.25M million detector simulation jobs were completed on
Titan, for a total of 225M events processed. This is equivalent to 9\% of all
the 250M detector simulations performed by ATLAS in the same period of time, and
3.5\% of the 6.6B events processed by those jobs. Comparatively, Titan
contributed 3.9\% of the total of around 200K cores available to ATLAS on Grid,
Cloud, and HPC infrastructures together. When accounting for the amount of
unused backfill availability and the rate of improvement of PanDA efficiency on
Titan, these figures confirms the relevance of supercomputers' resource
contribution to the LHC Run 3.

@@

Detector simulations executed on Titan, all process 100 events. This number has
been derived empirically by observing the average walltime of backfill
resources: too many events would \ldots\mtnote{I need to speak further to Sergey
to clarify this.}. Usually, detector simulations performed on the Grid process a
higher number of events than those performed on Titan. This explains the
difference between the percentage of detector simulations performed on Titan
(9\%) and of events computed on Titan for those simulations (3.9\%). PanDA Event
Service~\cite{panda-event-service} would address this limitation: each PanDA
Broker would be able to pull a dynamic number of events for each simulation,
depending on the available walltime. This functionality is currently being
implemented.

@@

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit; and (ii) the distribution of the time taken by the
Geant4 toolkit to simulate a certain number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries. Once initialized,
every algorithm and service of AthenaMP is configured by a set of Python
scripts. Both these operations result in a large number of read operations on
the filesystem, including those required to access of small python scripts.

@@

Initially, all shared libraries of AthenaMP and the configuration python scripts
were stored on the Spider 2 Lustre file system. However, the I/O patterns of the
initialization and configuration stages degraded the performance of the
filesystem (Figure~\ref{}). Since Spider 2 is a center-wide file system, this
resulted in lower overall interactive and metadata performance for all OLCF
resources and users. \mtnote{I am afraid the details about the trace are too
specific given the space constraints of a SC submission. Please feel free to
uncomment it if you disagree.}
% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.

@@

This issue was addressed by moving the AthenaMP distribution to a read-only NFS
directory made accessible from the Titan work nodes. This eliminated the problem
of metadata contention, improving metadata read performance from \~6,300 seconds
on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the input
files of each detector simulation job were stored in a ramdisk on the work
nodes. This offered a marked performance improvement: from 1,320 seconds on
Lustre to 40 seconds on NFS.

@@

At runtime, detector simulation jobs showed performance degradation when all the
16 cores of Titan's work nodes were used. As shows in Figure~\ref{}, the mean
runtime for 8 concurrent jobs computing 100 events is XXXs, with each job
executed in 8 distinct processes, one per core. The mean runtime for 16
concurrent jobs computing 100 events is instead XXXs. This can be explained by
looking at the architecture of the cache in the AMD Opteron 6274 CPUs. This
processor shares the level 1 instruction cache between couples of cores,
impacting the performance of \ldots computation.

The compute performance of Titan for detector simulation jobs is, on average,
lower that that of LCWG. Figure~\ref{} shows the difference in execution time of
the same type of simulation between Titan and \ldots. This is not an issue with
the design of PanDA Brokers and, more in general, with the migration of part of
the ATLAS workload to supercomputers: Titan has been active since 2012 and it is
to be expected that more modern CPUs offer better performance. A pilot project
is already in place to test PanDA Broker on Summit, the next generation
supercomputer that will become active in 2018.

@@

The current design and architecture of the PanDA Broker is proving to be as
reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
Brokers were responsible for around the 19\% of the failures, compared to the
29\% of failures produced by the JobDispatcher module of the PanDA Server, and
the 13\% failures produced by Geant4. The current failure rate confirms the
benefits given by reusing most of the code base of the PanDA Pilot for
implementing the PanDA Broker. It also shows that adopting third-party libraries
like RADICAL-SAGA did not have a measurable adverse effect on reliability.

Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
recorded via the PanDA Brokers between January 2016 and February 2017. The
initial preponderance of unknown failures was progressively addressed by a more
accurate failure mode developed for the PanDA Broker and Titan. It also shows
the progressive improvement of the broker's reliability: In Februrary 2017, most
of the failures were produced by the initialization and configuration phases of
AthenaMP or by Geant4 failed executions.\mtnote{If we want to keep this diagram,
we will need to plot it from the CSV to make it readable.}

>>>>

Currently, ATLAS deploys 20 instances of PanDA Broker on 4 Titan's DTNs, 5
instances per DTN. Each broker submits and manages the execution of 15 to 300
detector simulation jobs, one job for each work node, and a theoretical maximum
concurrent use of 96,000 cores. Since November 2015, PanDA Brokers have operated
only in backfill mode, without a defined time allocation, and running at the
lowest priority on Titan.

We evaluated the efficiency, scalability and reliability of the PanDA Brokers
along two dimensions: (i) the amount of backfill availability utilized by the
PanDA Brokers; and (ii) the amount of jobs walltime spent performing detector
simulations. We based these evaluations on measuring the number of cores
utilized by ATLAS on Titan, the overall backfill availability, the number of
detector simulations performed on all the resources available to ATLAS and on
Titan, and the failure rate of these simulations. All the measurements were
performed between January 2016 and February 2017.

@@

On February 2017, PanDA Brokers used almost twice as much backfill availability
than in any other month. No relevant code update was made during that period and
logs indicated that the brokers were able to respond more promptly to backfill
availability. This is likely due to hardware upgrades on the DTNs. The absence
of continuous monitoring of the nodes does not allow to quantify bottlenecks but
spot measurements of the DTNs load indicate that a faster CPU and better
networking were likely responsible for the improved performance.

Investigations showed an average CPU load of 3.6\% on the upgraded DTNs. As
such, further hardware upgrades seem unlikely to improve significantly the
performance of PanDA Brokers. Nonetheless, the current load suggests that the
number of brokers per DTN could be increased. This would enable the submission
of a larger number of concurrent jobs to Titan's PBS queue, allowing for PanDA
to consume a higher percentage of backfill availability.

Figure~\ref{fig:backfill-utilization} shows backfill utilization efficiency from
January 2016 to February 2017. Efficiency is defined as the fraction of Titan’s
cores available via backfill that was utilized by the PanDA Brokers. The brokers
reached 18\% average efficiency with a minimum 8.9\% efficiency on April 2016
and a maximum 33.5\% efficiency on February 2017. The number of total backfill
cores available was 38.1M in April 2016, and 33.1M in February 2017. This shows
that the measured difference in efficiency did not depend on a comparable
difference in total backfill availability.

@@

Figure~\ref{fig:hpc-workload-utilization} shows that between January 2016 and
February 2017, about 2.25M million detector simulation jobs were completed on
Titan, for a total of 225M events processed. This is equivalent to 9\% of all
the 250M detector simulations performed by ATLAS in the same period of time, and
3.5\% of the 6.6B events processed by those jobs. Comparatively, Titan
contributed 3.9\% of the total of around 200K cores available to ATLAS on Grid,
Cloud, and HPC infrastructures together. These figures confirms the relevance of
supercomputers' resource contribution to the LHC Run 3, especially when
accounting for the amount of unused backfill availability and the rate of
improvement of PanDA efficiency.

@@

All the detector simulations executed on Titan process 100 events. This number
has been chosen by observing the average walltime of backfill resources: too
many events would \ldots\mtnote{I need to speak further to Sergey to clarify
this.}.

Usually, detector simulations performed on the Grid process a higher number of
events than those on Titan. This explains the difference between the percentage
of detector simulations (9\%) and of events computed by those simulations
(3.9\%) on Titan. PanDA Event Service~\cite{panda-event-service} will contribute
to address this limitation: each PanDA Broker will be able to pull a variable
number of events for each simulation, depending on the available walltime. This
functionality is currently being implemented.

@@

We use two main parameters to measure the performance of the detector simulation
jobs submitted to Titan: (i) the time taken to setup
AthenaMP~\cite{aad2010atlas}, the ATLAS software framework integrating the
GEANT4 simulation toolkit~\cite{agostinelli2003geant4}; and (ii) the
distribution of the time taken by the Geant4 toolkit to simulate a certain
number of events.

AthenaMP has an initialization and configuration stage. At initialization time,
AthenaMP is assembled from a large number of shared libraries, depending on the
type of payload that will have to be computed. Once initialized, every algorithm
and service of AthenaMP is configured by a set of Python scripts. Both these
operations result in a large number of read operations on the filesystem,
including those required to access of small python scripts.

@@

Initially, all the shared libraries of AthenaMP and the python scripts for the
configuration stage were stored on the Spider 2 Lustre file system. However, the
I/O patterns of the initialization and configuration stages degraded the
performance of the filesystem (Figure~\ref{}\mtnote{We may want to aggregate
some of the diagrams about lustre's experiments here}). Since Spider 2 is a
center-wide file system, this resulted in  performance degradation for all OLCF
resources and users. \mtnote{I am afraid the details about the trace are too
specific given the space constraints of a SC submission. Please feel free to
uncomment it if you disagree.}
% As can be seen in Listing~\ref{mdstrace} the metadata I/O activity for ATLAS
% exhibits a spike corresponding to the beginning of the runs before tapering
% off.

@@

After careful characterization of the impact of ATLAS jobs on Lustre, the
performance degradation was addressed by moving the AthenaMP distribution to a
read-only NFS directory, shared among DTNs and work nodes. This eliminated the
problem of metadata contention, improving metadata read performance from \~6,300
seconds on Lustre to \~1,500 seconds on NFS. Further, at configuration time, the
input files of each detector simulation job were stored to a ramdisk on the work
nodes. This offered a marked improvement of reading time: from 1,320 seconds on
Lustre to 40 seconds on NFS.

@@

At runtime, detector simulation jobs showed performance degradation when all the
16 cores of Titan's work nodes were used. As shows in
Figure~\ref{}\mtnote{Sergery: do you have a diagram showing the difference in
performance between using 16 and 8 cores for a detector simulation job with 100
events?}, the mean runtime for 8 concurrent jobs computing 100 events is XXXs,
with each job executed in 8 distinct processes, one per core. The mean runtime
for 16 concurrent jobs computing 100 events is instead XXXs. This can be
explained by looking at the cache architecture of the AMD Opteron 6274 CPU. In
this processor, every two cores share the level 1 instruction cache, impacting
the performance of \ldots\mtnote{Do we kow what instruction set is affected?}
computation.

The compute performance of Titan for detector simulation jobs is, on average,
lower that that of LCWG. Figure~\ref{}\mtnote{Sergey: I remember you showed us
some plots comparing Geant4 simulation's performance on Titan and some of the
Grid sites. Could we use one or two of the here?} shows the difference in
execution time of the same type of simulation between Titan and
\ldots\mtnote{Add the name of the grid sites here}. This is not an issue with
the design of PanDA Brokers and, more in general, with the migration of part of
the ATLAS workload to supercomputers: Titan has been active since 2012 and more
modern CPUs are expected to offer better performance. A pilot project is already
in place to test PanDA Broker on Summit, the next generation supercomputer that
will become active in 2018 at OLCF.

@@

The current design and architecture of the PanDA Broker is proving to be as
reliable as PanDA Pilot when used on the WLCG. Between Jan 2016 and Feb 2017,
the overall failure rate of all the ATLAS detector simulation jobs was 14\%,
while the failure rate of jobs submitted to Titan was a comparable 13.6\%. PanDA
Brokers were responsible for around the 19\% of the failures, compared to the
29\% of failures produced by the JobDispatcher module of the PanDA Server, and
the 13\% failures produced by the Geant4 toolkit. The current failure rate of
the PanDA Brokers confirms the benefits of reusing most of the code base of the
PanDA Pilot for implementing the PanDA Broker. It also shows that adopting
third-party libraries like RADICAL-SAGA did not have a measurable adverse effect
on reliability.

Figure~\ref{fig:failures-titan} shows a breakdown of the type of failures
recorded by the PanDA Brokers between January 2016 and February 2017. The
initial preponderance of unknown failures was progressively addressed by a more
accurate failure model, specifically tailored to the PanDA Broker and Titan. It
also shows the progressive improvement of the broker's reliability: In Februrary
2017, most of the failures were produced by the initialization and configuration
phases of AthenaMP or by failed executions of the payload.\mtnote{If we want to
keep this diagram, we will need to plot it from the CSV to make it readable.}
